---
title: "R Notebook"
output: html_notebook
---

```{r}
library(tidyverse)
library(readr)
```


```{r}
#candy_rating_analysis1 <- read_csv("../clean_data/final_candy_ratings.csv", col_types = cols(`Internal ID` = col_number()))

```

# Loading the library`

```{r}
library(tidyverse)
library(readr)
```

# Reading the dataset and passing the data types for the column`

```{r}
candy_rating_analysis1 <- read_csv("../clean_data/final_candy_ratings.csv", col_types = cols(`Internal ID` = col_integer(),
  gender = col_character(),
  country = col_character(),
  state_county_province = col_character(),
  date = col_date()))
```

# What is the total number of candy ratings given across the three years. (number of candy ratings, not number of raters. Don’t count missing values)`

```{r}

total <- candy_rating_analysis1 %>% 
 count(reaction) 
sum(total$n[1:3])
```


# What was the average age of people who are going out trick or treating and the average age of people not going trick or treating?`

```{r}

candy_rating_analysis1 %>% 
  select(going_trick_treat, age) %>% 
  group_by(going_trick_treat) %>% 
  summarise(avg = mean(age, na.rm = TRUE))

```


# For each of joy, despair and meh, which candy bar received the most of these ratings?`

```{r}

##JOY##

j <- candy_rating_analysis1 %>% 
  select(candy_category, reaction) %>% 
  group_by(candy_category) %>% 
  filter(reaction == "JOY") %>% 
 summarise(joy = n()) %>% 
  arrange(desc(joy)) %>% 
  head(1)


###DESPAIR###

d <- candy_rating_analysis1 %>% 
  select(candy_category, reaction) %>% 
  group_by(candy_category) %>% 
  filter(reaction == "DESPAIR") %>% 
 summarise(despair = n()) %>% 
  arrange(desc(despair)) %>% 
  head(1)


###MEH##

m <- candy_rating_analysis1 %>% 
  select(candy_category, reaction) %>% 
  group_by(candy_category) %>% 
  filter(reaction == "MEH") %>% 
 summarise(meh = n()) %>% 
  arrange(desc(meh)) %>% 
  head(1)


mj <-full_join(m,j)
jdm <- full_join(mj,d, by = "candy_category")

jdm
```

# How many people rated Starburst as despair?`

```{r}

candy_rating_analysis1 %>% 
  select(candy_category, reaction) %>% 
  filter(candy_category == "Starburst" & reaction == "DESPAIR") %>% 
  count()
```
# For the next three questions, count despair as -1, joy as +1 and meh as 0.`

```{r}

candy_rating_analysis2 <- candy_rating_analysis1 %>%
  #select(candy_category, reaction) %>% 
  mutate(reaction_score = case_when(
                                    reaction == "JOY" ~ 1,
                                    reaction == "DESPAIR" ~-1,
                                    reaction == "MEH" ~ 0
  ),.after = reaction)
candy_rating_analysis2
```

# What was the most popular candy bar in each year?`

```{r}

candy_rating_analysis2 %>% 
  mutate(date = as.numeric(format(date,'%Y'))) %>% 
  group_by(date, candy_category) %>% 
  summarise(popular = mean(reaction_score, na.rm = TRUE)) %>% 
  slice_max(popular, n=1) %>% 
  arrange(desc(popular))

```


# What was the most popular candy bar by this rating for people in US, Canada, UK and all other countries?`


```{r}

candy_rating_analysis2 %>% 
  group_by(country, candy_category) %>% 
  summarise(popular = mean(reaction_score, na.rm = TRUE)) %>% 
  filter(country == "USA"| country == "UK" | country == "Canada") %>% 
  slice_max(popular, n=1) %>% 
  arrange(desc(popular))


```


# What was the most popular candy bar in each year?`

```{r}

candy_rating_analysis2 %>% 
  mutate(new = as.numeric(format(date,'%Y'))) %>% 
  distinct(new)

candy_rating_analysis2 %>% 
  distinct (date)


candy_rating_analysis2 %>% 
  mutate(date = as.numeric(format(date,'%Y'))) %>% 
  group_by(date, candy_category) %>% 
  summarise(popular = mean(reaction_score, na.rm = TRUE)) %>% 
  slice_max(popular, n=1) %>% 
  arrange(desc(popular))
```


# What was the most popular candy bar by this rating system for each gender in the dataset?`

```{r}

candy_rating_analysis2 %>% 
  group_by(gender, candy_category) %>% 
  summarise(popular = mean(reaction_score, na.rm = TRUE)) %>% 
  slice_max(popular, n=1) %>% 
  arrange(desc(popular))
 


```






