---
title: "dirty_data_project_task4"
author: "Monica"
date: "22/04/2021"
output: html_document
---

## INTRODUCTION 
### *CANDY_RANKING_DATA* DATASET

The dataset for the the dirty project consists of 3 files excel files. Each file
had the data for the year 2015, 2016, 2017 respectively.The three datasets 
had columns which were either present in all the three dataset or just one dataset. 
However, at any given point atleast one column was common amongst any two dataset. 
The motive of this project is to do date cleaning and data wrangling. 
Below are the steps taken to present the final dataset in a form that is tidy and 
the user can do analysis of the candy ratings. The records mainly contains a reaction 
on different kinds of candies. The customer is of various age group and also from
different countries. The reaction is either *JOY* or *DESPAIR* or *MEH*


## ASSUMPTIONS

#### boing_boing_candy_dataset_2015

1. The timestamp needs to be divided into 2 columns namely - date and time.
2. The name of the column needs to be small and readable
3. Irrelevant columns/rows which doesn't have any information regarding the candy
   ratings should be removed completely from the table.
4. Column age which has information other than number should be replaced with NA. Also,
   values in the age column should be between 0 and 100.
5. Any reaction other than *JOY* or *DESPAIR* should be NA.


#### boing_boing_candy_dataset_2016

1. The timestamp needs to be divided into 2 columns namely - date and time.
2. The name of the column needs to be small and readable
3. Irrelevant columns/rows which doesn't have any information regarding the candy
   ratings should be removed completely from the table.
4. Column age which has information other than number should be replaced with NA. Also,
   values in the age column should be between 0 and 100.
5. Any reaction other than *JOY* or *DESPAIR* or *MEH* should be NA. 
6. Country column should be in the same case for the values which means the same.
7. Shortforms of certain province/county should be put as full name.


#### boing_boing_candy_dataset_2017

1. The timestamp needs to be divided into 2 columns namely - date and time.
2. The name of the column needs to be small and readable
3. Irrelevant columns/rows which doesn't have any information regarding the candy
   ratings should be removed completely from the table.
4. Column age which has information other than number should be replaced with NA. Also,
   values in the age column should be between 0 and 100.
5. Any reaction other than *JOY* or *DESPAIR* should be NA. 
6. Country column should be in the same case for the values which means the same.
7. Shortforms of certain province/county should be put as full name.



## DATA CLEANING 

The first step of cleaning the data is to read the data. Depending upon the file 
type, you can use the function. For the function to run, you would need to install 
package and load library.

Across all the dataset, there were couple of same cleaning pattern followed. 
These steps are mentioned below:- 

`Separate column *Timestamp* into *date* and *time*`

```{r eval = FALSE}
data_boing_candy_2015v1 <- data_boing_candy_2015 %>% 
  separate(Timestamp, c("date", "time"), sep = " " )
class(data_boing_candy_2015v1)
```


`Changing column names`

```{r eval = FALSE}

names(data_boing_candy_2015v1)[names(data_boing_candy_2015v1) == "How old are you?"] <- "age"
names(data_boing_candy_2015v1)[names(data_boing_candy_2015v1) == "Are you going actually going trick or treating yourself?"] <- "going_trick_treat"
names(data_boing_candy_2015v1)[names(data_boing_candy_2015v1) == "Please leave any remarks or comments regarding your choices."] <- "comments"

```


`Removing irrelevant columns which will not affect the candy ratings`

```{r eval = FALSE}
data_boing_candy_2016v1$`Please list any items not included above that give you JOY.` <- NULL
data_boing_candy_2016v1$`[York Peppermint Patties] Ignore` <- NULL
data_boing_candy_2016v1$`Please leave any witty, snarky or thoughtful remarks or comments regarding your choices.` <- NULL
data_boing_candy_2016v1$`Guess the number of mints in my hand.` <- NULL
data_boing_candy_2016v1$`Betty or Veronica?` <- NULL
```

`Replacing the unacceptable value of age with NA`

```{r eval = FALSE}
options(scipen = 9999)
data_boing_candy_2016v1$age <- as.integer(as.character(data_boing_candy_2016v1$age))
```

`Making table longer to have minimum column and compact table`

```{r eval = FALSE}
data_boing_candy_2016v2 <- data_boing_candy_2016v1 %>%
  pivot_longer(cols = `[100 Grand Bar]`:`[York Peppermint Patties]`,
               names_to = "candy_category", 
               values_to = "reaction") 
```


`Tidying of the column country - putting same values that meant same. Representing`
`only few lines code `

```{r eval = FALSE}

data_boing_candy_2016v3 <- data_boing_candy_2016v2 %>% 
  mutate(country = recode(country, "usa" = "USA", "us" = "USA", "US" =  "USA", "Usa" = "USA",
                          "America" = "USA", "United States of America" = "USA",
                          "United States" = "USA", "USA!!!!!!" = "USA", 
                          "uSA" = "USA", "united states" = "USA", 
                          "USA USA USA USA" = "USA", "america" = "USA",
                          "netherlands" = "Netherlands", "kenya" = "Kenya", 
                          "USA (I think but it's an election year so who can really tell)" = "USA",
                          "USA!" = "USA", "Units States" = "Units States", "U.S." = "USA",
                          "United states" = "USA", "USA USA USA" = "USA", "USA! USA! USA!" = "USA",
                          "United Kingdom" = "UK", "United Kindom" = "UK", "Units States" = "USA", 
                          "the best one - usa" = "USA",
                          ))

```


`tidying the state_county_province column by replacing short names with the full names`

```{r eval = FALSE}
data_boing_candy_2016v4 <- data_boing_candy_2016v3 %>% 
  mutate(state_county_province = recode(state_county_province, "il" = "Illinois",
                                        "PA" = "Pennsylvania", "IN" = "Indiana",
                                        "the Mitten" = "Michigan"))
```

`Replacing "NA" with NA. The unwanted column values were initially recoded to string NA.`

```{r eval = FALSE}
data_boing_candy_2016v5  <- data_boing_candy_2016v4 %>% 
  mutate(country = replace_na("NA"))
```

`Replacing the string which contains punctuations. The values in the candy_category`
`were encapsulated with square bracket which had to be cleaned`

```{r eval = FALSE}
data_boing_candy_2016v6 <- data_boing_candy_2016v5 %>% 
  mutate(candy_category = str_replace_all(candy_category, "[[:punct:]]", ""))
```


`This code will accept age between 0 and 100`

```{r eval = FALSE}
data_boing_candy_2016v7 %>% 
  mutate(age = if_else(age < 0 | age > 100, NA_integer_, age)) %>% 
  filter(age>100)

```


## DATA MERGING

`The three clean datasets were merged into one dataset which will be used for` 
`analysis purpose. This *final_candy_ratings* was then written as csv file. This`
`file is stored in the *clean_data* folder`

```{r eval = FALSE}

data_2015_2017 <- bind_rows(data_boing_candy_2015v5,data_boing_candy_2017v5)

data_2015_2017_2016 <- bind_rows(data_2015_2017,data_boing_candy_2016v7)

final_candy_ratings <- write_csv(data_2015_2017_2016,"final_candy_ratings.csv")

```



## TASK QUESTION AND ANALYSIS

Once the cleaned data is ready, we can read the dataset and start analysing the 
values. Here are few questions and the code.

`Loading the library`

```{r eval = FALSE}
library(tidyverse)
library(readr)
```

`Reading the dataset and passing the data types for the column`

```{r eval = FALSE}
candy_rating_analysis1 <- read_csv("../clean_data/final_candy_ratings.csv", col_types = cols(`Internal ID` = col_integer(),
  gender = col_character(),
  country = col_character(),
  state_county_province = col_character(),
  date = col_date()))
```

`What is the total number of candy ratings given across the three years. (number of candy ratings, not number of raters. Don’t count missing values)`

```{r eval = FALSE}

total <- candy_rating_analysis1 %>% 
 count(reaction) 
sum(total$n[1:3])
```


`What was the average age of people who are going out trick or treating and the average age of people not going trick or treating?`

```{r eval = FALSE}

candy_rating_analysis1 %>% 
  select(going_trick_treat, age) %>% 
  group_by(going_trick_treat) %>% 
  summarise(avg = mean(age, na.rm = TRUE))

```


`For each of joy, despair and meh, which candy bar received the most of these ratings?`

```{r eval = FALSE}

##JOY##

j <- candy_rating_analysis1 %>% 
  select(candy_category, reaction) %>% 
  group_by(candy_category) %>% 
  filter(reaction == "JOY") %>% 
 summarise(joy = n()) %>% 
  arrange(desc(joy)) %>% 
  head(1)


###DESPAIR###

d <- candy_rating_analysis1 %>% 
  select(candy_category, reaction) %>% 
  group_by(candy_category) %>% 
  filter(reaction == "DESPAIR") %>% 
 summarise(despair = n()) %>% 
  arrange(desc(despair)) %>% 
  head(1)


###MEH##

m <- candy_rating_analysis1 %>% 
  select(candy_category, reaction) %>% 
  group_by(candy_category) %>% 
  filter(reaction == "MEH") %>% 
 summarise(meh = n()) %>% 
  arrange(desc(meh)) %>% 
  head(1)


mj <-full_join(m,j)
jdm <- full_join(mj,d, by = "candy_category")

jdm
```

`How many people rated Starburst as despair?`

```{r eval = FALSE}

candy_rating_analysis1 %>% 
  select(candy_category, reaction) %>% 
  filter(candy_category == "Starburst" & reaction == "DESPAIR") %>% 
  count()
```


`What was the most popular candy bar in each year?`
```{r eval = FALSE}

candy_rating_analysis2 %>% 
  mutate(date = as.numeric(format(date,'%Y'))) %>% 
  group_by(date, candy_category) %>% 
  summarise(popular = mean(reaction_score, na.rm = TRUE)) %>% 
  slice_max(popular, n=1) %>% 
  arrange(desc(popular))

```


`For the next three questions, count despair as -1, joy as +1 and meh as 0.`

```{r eval = FALSE}

candy_rating_analysis2 <- candy_rating_analysis1 %>%
  #select(candy_category, reaction) %>% 
  mutate(reaction_score = case_when(
                                    reaction == "JOY" ~ 1,
                                    reaction == "DESPAIR" ~-1,
                                    reaction == "MEH" ~ 0
  ),.after = reaction)
candy_rating_analysis2
```


`What was the most popular candy bar by this rating for people in US, Canada, UK and all other countries?`


```{r eval = FALSE}

candy_rating_analysis2 %>% 
  group_by(country, candy_category) %>% 
  summarise(popular = mean(reaction_score, na.rm = TRUE)) %>% 
  filter(country == "USA"| country == "UK" | country == "Canada") %>% 
  slice_max(popular, n=1) %>% 
  arrange(desc(popular))


```


`What was the most popular candy bar in each year?`

```{r eval = FALSE}

candy_rating_analysis2 %>% 
  mutate(new = as.numeric(format(date,'%Y'))) %>% 
  distinct(new)

candy_rating_analysis2 %>% 
  distinct (date)


candy_rating_analysis2 %>% 
  mutate(date = as.numeric(format(date,'%Y'))) %>% 
  group_by(date, candy_category) %>% 
  summarise(popular = mean(reaction_score, na.rm = TRUE)) %>% 
  slice_max(popular, n=1) %>% 
  arrange(desc(popular))
```


`What was the most popular candy bar by this rating system for each gender in the dataset?`

```{r eval = FALSE}

candy_rating_analysis2 %>% 
  group_by(gender, candy_category) %>% 
  summarise(popular = mean(reaction_score, na.rm = TRUE)) %>% 
  slice_max(popular, n=1) %>% 
  arrange(desc(popular))
 


```


