---
title: "task1_markdown"
author: "Monica"
date: "22/04/2021"
output: html_document
---

## INTRODUCTION 
### *Decathlon Data* DATASET

The dataset *Decathalon_data* gives the information about the time taken by different
player in different sports discipline. The time mentioned is in seconds and milliseconds.


## ASSUMPTIONS

None


## DATA CLEANING 

The first step of cleaning the data is to read the data. Depending upon the file 
type, you can use the function. For the function to run, you would need to install 
package and load library.

`Reading the data`

```{r eval = FALSE}
data_clean_decathalon_v1 <- read_rds("raw_data/decathlon.rds")

```


`Getting to know the dataset by passing few functions`

```{r eval = FALSE}

view(data_clean_decathalon_v1)

#number of rows
nrow(data_clean_decathalon_v1)

#number of columns
ncol(data_clean_decathalon_v1)

#row and column 
dim(data_clean_decathalon_v1)

#names of the column
names(data_clean_decathalon_v1)

#view of the data type
glimpse(data_clean_decathalon_v1)

#view summary
summary(data_clean_decathalon_v1)

#viewing the class
class(data_clean_decathalon_v1)

```

## DATA CLEANING

To use the data cleaning we would need to load the package *JANITOR* so that we 
use functions that will help tidy the dataset.

`load library`

```{r eval = FALSE}
library (janitor)
```

`cleaning dataset using functions from janitor package`

```{r eval = FALSE}

data_clean_decathalon_v2 <- data_clean_decathalon_v1 %>% 
clean_names() %>% 
rownames_to_column("names") %>% 
mutate(names = str_to_sentence(names)) %>% 
view(data_clean_decathalon_v2)
glimpse(data_clean_decathalon_v2)
```


`Renaming the column name so that the column will be tidied`

```{r eval = FALSE}
names(data_clean_decathalon_v2)[names(data_clean_decathalon_v2) == "x100m"] <- "100_m"
names(data_clean_decathalon_v2)[names(data_clean_decathalon_v2) == "x110m_hurdle"] <- "110_m_hurdle"
names(data_clean_decathalon_v2)[names(data_clean_decathalon_v2) == "x400m"] <- "400_m"
names(data_clean_decathalon_v2)[names(data_clean_decathalon_v2) == "x1500m"] <- "1500_m"
names(data_clean_decathalon_v2)
```


`Writing data to csv file`

```{r eval = FALSE}
write_csv(data_clean_decathalon_v2, "decathalon_cleaned_data.csv")
```


## DATA WRANGLING

`Loading the library`

```{r eval = FALSE}
library(tidyverse)
library(readr)
```

`Reading the dataset using read function`

```{r eval = FALSE}
decat_analysis_datav1 <- read_csv("decathalon_cleaned_data.csv")
decat_analysis_datav1
```

`Who had the longest long jump seen in the data?`

```{r eval = FALSE}
decat_analysis_datav1 %>% 
  select(names, long_jump) %>% 
  arrange(desc(long_jump)) %>% 
  head(1)
  
```

`What was the average 100m time in each competition?`

```{r eval = FALSE}

decat_analysis_datav1 %>% 
  select (`100_m`, competition) %>% 
  group_by(competition) %>% 
  summarise(avg = mean(`100_m`))
 
```

`Who had the highest total points across both competitions?`

```{r eval = FALSE}
decat_analysis_datav1 %>% 
  select (names,competition, points) %>% 
  group_by(competition) %>% 
  slice_max(points, n = 1)
  
```


`What was the shot-put scores for the top three competitors in each competition?`

```{r eval = FALSE}
decat_analysis_datav1 %>% 
  select (names,competition, shot_put) %>% 
  group_by(competition) %>% 
  slice_max(shot_put, n = 3)
```

`What was the average points for competitors who ran the 400m in less than 50 seconds vs. those than ran 400m in more than 50 seconds?`

```{r eval = FALSE}
more_than_50 <- decat_analysis_datav1 %>% 
  select(names, `400_m`, competition) %>% 
  filter (`400_m` > 50.00) %>% 
  group_by(competition) %>% 
  summarise(avg_more_than_50 = mean(`400_m`))
  
more_than_50

less_than_50 <- decat_analysis_datav1 %>% 
  select(names, `400_m`, competition) %>% 
  filter (`400_m` < 50.00) %>% 
  group_by(competition) %>% 
  summarise(avg_less_than_50 = mean(`400_m`))

less_than_50

full_join(more_than_50, less_than_50)
```




